\documentclass{article}

\begin{document}

{\bf Re: Strengthening Research and Teaching.}

Provost Christ identifies our two main products as a university ---
teaching and research. Since faculty time and attention are finite
and precious resources, there is an underlying tradeoff between
these.  Given the desirability of increasing research revenues,
without compromising teaching standards, we have to focus on methods
of decreasing faculty overheads in these activities. Luckily our
current situation provides with some low hanging fruit --- the lack
of teaching support, including insufficient TAS budgets, is a major
drain on faculty time. Back of the envelope calculations suggest
that the tradeoff would be extremely favorable.


{\bf Re: Broadening Program Evaluation.}

Provost Christ brings up the need to focus on market and margin 
in addition to mission for our university. As she rightly points
out, faculty have traditionally focussed largely on mission, and
changing this mission oriented culture has serious risk, since 
our faulty are performing extremely well by all metrics. Market 
and means can much more readily be addressed by the other leg of the
university - the administration - without compromising their mission. 
Provost Christ's directive that every dollar counts in conjunction
with an analysis of UCB versus UCSB on effective
overhead rates suggests that we have significant opportunities for 
greater efficiency, without compromising mission, on the administrative side.   

\end{document}
